Refunctionalization of Abstract Abstract Machines (Functional Pearl)

### 2018-05-31 paper/paper.tex
reformat

  In this paper, we show their functional correspondence and how to syntactically 
  transform small-step abstract abstract machines into big-step abstract definitional
  interpreters.
  The transformations include linearization, fusing, disentangling, refunctionalizing,
  and un-CPS to direct-style with delimited controls. Linearization expresses
  non-deterministic choices by first-order data types, after which refunctionalization
  sequentializes the evaluation order by higher-order functions.
  All transformations properly handle the collecting semantics and the
  nondeterminism of abstract interpretation.
  %After each transformation, we also obtain an intermediate form of abstract interpreter.

  Following the idea that in deterministic languages,
  evaluation contexts of reduction semantics are defunctionalized continuations,
  we further show that in nondeterministic languages, evaluation contexts are
  refunctionalized to extended continuations style.
  Remarkably, we reveal how precise call/return matching in control-flow
  analysis is obtained by refunctionalizing a small-step abstract
  abstract machine with proper caching.

  %As well as we explain how it been lost by defunctionalizing the higher-order continuations into an first-order data type.
\end{abstract}

@@

\item The last transformation we show is that of transforming the refunctionalized AAM
  to a \textbf{direct-style} interpreter (Section~\ref{directstyle}) by using delimited controls.

@@

\item In the first @case@ statement shown in the previous code,
the control of the current state matches as a @Let@ expression, with its right-hand side a
function application.
By calling the @atomicEval@ evaluator, we obtain the closure for which the callee @f@ stands.
The successor state's control then transfers to the @body@ expression of the closure
with an updated environment and store. The new environment is extended
from the closure's environment and mapped from @v@ to a fresh address @addr@.
The new store is extended with @addr@ mapping to the value of @ae@,
which in turn is evaluated from @atomicEval@.
Finally, a new frame @frame@ is pushed onto the stack @k@, where
@frame@ contains the variable name @x@ at the left-hand position of the @Let@,
the body expression of @Let@, and the lexical environment of the body expression.

@@

Similar to the CESK machines, to run (analyze) a program
we first use the @inject@ function to construct the initial state given to the program.
Note that the initial continuation store has a built-in mapping that maps the continuation
address for @Halt@ to an empty set of continuations.
We also provide an empty program contour as our initial time.
\begin{lstlisting}
def inject(e: Expr): State =
  State(e, Map(),
        Store[BAddr, Storable](Map()),
        Store[KAddr, Cont](Map(Halt -> Set())),
        Halt, List())
\end{lstlisting}

@@

The above code shows the skeleton of @drive_step@.
At the end of the first case of pattern matching, we invoke @drive_step@ recursively\footnote{The code for constructing new environments, stores, and frames is elided for simplicity of presentation.}.
The code for the second case is replaced entirely by a function call to @continue@.
The astute reader will notice a vague shape of interpreter in continuation style has emerged.

>>>>

  In this paper, we show their functional correspondence and how to syntactically 
  transform small-step abstract abstract machines into big-step abstract  definitional
  interpreters.
  The transformations include linearization, fusing, disentangling, refunctionalizing,
  and un-CPS to direct-style with delimited controls. Linearization expresses
  non-deterministic choices by first-order data types, after which refunctionalization
  sequentializes the evaluation order by higher-order functions.
  All transformations properly handle the collecting semantics and the nondeterminism 
  of abstract interpretation.

  Following the idea that in deterministic languages, evaluation contexts of reduction
  semantics are defunctionalized continuations, we further show that in nondeterministic 
  languages, evaluation contexts are refunctionalized to extended continuations style.
  Remarkably, we reveal how precise call/return matching in control-flow analysis is 
  obtained by refunctionalizing a small-step abstract abstract machine with proper caching.
\end{abstract}

@@

  \item The last transformation we show is that of transforming the refunctionalized AAM to a
    \textbf{direct-style} interpreter (Section~\ref{directstyle}) by using delimited controls.
\end{itemize}

@@

\item In the first @case@ statement shown in the previous code, the control of the current 
  state matches as a @Let@ expression, with its right-hand side a function application.
  By calling the @atomicEval@ evaluator, we obtain the closure for which the callee @f@ stands.
  The successor state's control then transfers to the @body@ expression of the closure with 
  an updated environment and store. The new environment is extended from the closure's 
  environment and mapped from @v@ to a fresh address @addr@. The new store is extended with
  @addr@ mapping to the value of @ae@, which in turn is evaluated from @atomicEval@.
  Finally, a new frame @frame@ is pushed onto the stack @k@, where the @frame@ contains 
  the variable name @x@ at the left-hand position of the @Let@, the body expression of 
  @Let@, and the lexical environment of the body expression.

@@

Similar to the CESK machines, to run (analyze) a program we first use the @inject@ function
to construct the initial state given to the program. Note that the initial continuation store
has a built-in mapping that maps the continuation address for @Halt@ to an empty set of 
continuations. We also provide an empty program contour as our initial time.

\begin{lstlisting}
def inject(e: Expr): State =
  State(e, Map(),
        Store[BAddr, Storable](Map()),
        Store[KAddr, Cont](Map(Halt -> Set())),
        Halt, List())
\end{lstlisting}

@@

The above code shows the skeleton of @drive_step@.
At the end of the first case of pattern matching, we invoke @drive_step@ recursively 
\footnote{The code for constructing new environments, stores, and frames is elided for 
simplicity of presentation.}.
The code for the second case is replaced entirely by a function call to @continue@.
The astute reader will notice a vague shape of interpreter in continuation style has emerged.
